---
title: "EDA"
output: html_notebook
---

```{r}
library(readr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(GGally)
```

```{r message = FALSE}
df <- read_csv("cleaned_data.csv")
```

### questions to answer
1. What is his win percentage overall, grouped by color?
2. Which is his best opening overall, grouped by color?
3. Which opening does he play the most overall, grouped by color?
4. Does he lose more against higher rated opponents?
5. What is the average move count in his games?
6. Which year did he perform the best in?
7. Who does he play with most often?

### what is his win percentage overall & grouped by color?
As white, H.N wins more, loses & draws less than as black. As black
he loses slightly more often than wins.
```{r include=F}
overall_win <- df %>%
  group_by(Result) %>%
  summarise(count = n()) %>%
  mutate(Percentage = count / sum(count) * 100) %>%
  select(-count) 
overall_win
 
overall_win %>% ggplot(aes(x = Result, y = Percentage)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(title = "Ratio of Wins, Losses & Draws overall")


winrate <- df %>%
  group_by(Color, Result) %>%
  summarise(count = n()) %>%
  mutate(Percentage = count / sum(count) * 100) %>%
  select(-count) 
winrate

winrate %>% ggplot(aes(x = Color, y = Percentage, fill = Result)) +
  geom_bar(stat = "identity", width = 0.6) +
  coord_flip() +
  labs(title = "Ratio of Wins, Losses & Draws by Color")
```

### which openings has higher winrates overall & grouped by color?
```{r}
overall_op <- df %>%
  group_by(Opening, Result)
  


```
