---
title: "EDA"
output: html_notebook
---

### Load libraries
```{r}
library(readr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(gridExtra)
```

### Import and view data
```{r message = FALSE}
df <- read_csv("cleaned_data.csv")
head(df)
```

#### questions to answer
1. *What is his win percentage overall, grouped by color?*
2. Which is his best opening overall, grouped by color?
3. *Which opening does he play the most overall, grouped by color?*
4. *Does he lose more against higher rated opponents?*
5. *What is the average move count in his games?*
6. *Which year did he perform the best in?*
7. *Who does he play with most often?*

### What is his win percentage overall?
```{r echo=F, message = FALSE}
# ratio of win, draw, loss
overall_win <- df %>%
  group_by(Result) %>%
  summarise(count = n()) %>%
  mutate(Percentage = round(count / sum(count) * 100, 2)) %>%
  select(-count) 

overall_win %>%
  ggplot(aes(x = reorder(Result, -Percentage), y = Percentage)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    title = "Ratio of Wins, Losses & Draws overall",
    x = "Result"
  )
   
# ratio of win, draw, loss by color
winrate <- df %>%
  group_by(Color, Result) %>%
  summarise(count = n()) %>%
  mutate(Percentage = count / sum(count) * 100) %>%
  select(-count) 

winrate %>% ggplot(aes(x = Color, y = Percentage, fill = Result)) +
  geom_bar(stat = "identity",position = "fill", width = 0.6) +
  coord_flip() +
  labs(
    title = "Ratio of Wins, Losses & Draws by Color",
    y = "Win:Loss:Draw"
  )
```

### Which opening does he play the most overall & grouped by color?
```{r echo = FALSE, message = FALSE}
# openings and their game count
overall_op <- df %>%
  group_by(Opening) %>%
  summarise(N_games = n()) %>%
  arrange(desc(N_games))

# top 10 most played openings
head(overall_op, 10) %>%
  ggplot(aes(x = reorder(Opening, N_games), y = N_games)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Most Played Openings by Hikaru Nakamura",
    x = "Opening",
    y = "Number of Games"
  )

# top 10 most played openings grouped by color
top_by_color <- df %>%
  group_by(Color, Opening) %>%
  summarise(N_games = n()) %>%
  top_n(n = 10, wt = N_games)

# top 10 most played openings by white
white <- head(top_by_color, 10) %>%
  ggplot(aes(x = reorder(Opening, N_games), y = N_games)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Most played openings as white",
    x = "Openings",
    y = "Number of Games"
  )

# top 10 most played openings by black
black <- tail(top_by_color, 10) %>%
  ggplot(aes(x = reorder(Opening, N_games), y = N_games)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Most played openings as black",
    x = "Openings",
    y = "Number of Games"
  )

grid.arrange(white, black)
```


### Does he lose more against higher rated opponents?
```{r echo = FALSE, message = FALSE}
# results when playing against/not against higher rated opponents
vs_high <- df %>%
  mutate(Vs_higher = if_else(Opponent_rating > Rating, "yes","no")) %>%
  group_by(Vs_higher, Result) %>%
  summarise(count = n()) %>%
  mutate(Percentage = count / sum(count) * 100) %>%
  select(-count) %>%
  arrange(desc(Vs_higher)) %>%
  ggplot(aes(x = Vs_higher, y = Percentage, fill = Result)) +
  geom_bar(stat = "identity", width = 0.6) +
  coord_flip() +
  labs(
    title = "Ratio of Wins, Losses & Draws against higher/lower opponents",
    x = "VS higher-rated opponents?",
    y = "Win:Loss:Draw"
  )
```

### What is the average move count in his games & does move count affect the result of games?
```{r echo = FALSE, message = FALSE}
# move count distribution
df %>%
  ggplot(aes(x = Moves)) +
  geom_histogram(binwidth = 5) +
  labs(
    title = "distribution of move count of games",
    x = "move count of games"
  )

# avg moves for win/draw/loss
df %>%
  group_by(Result) %>%
  summarise(Mean_moves = mean(Moves)) %>%
  ggplot(aes(x = reorder(Result, Mean_moves), y = Mean_moves)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(
    title = "Average move counts of different game results",
    x = "Results",
    y = "Average move count"
  )
```

### Which year did he perform the best in?
ratings start from 2003, the year he became a gm. his highest rating was
in 2015.
```{r echo = FALSE, message = FALSE}
# highest rating of each year 2003-2020
df %>%
  group_by(Year) %>%
  filter(Year > 2002) %>%
  summarise(highest_rt = max(Rating, na.rm = TRUE)) %>%
  ggplot(aes(x = Year, y = highest_rt)) +
  geom_line() +
  scale_x_continuous(breaks = seq(2002, 2020, 2)) +
  labs(
    title = "Yearly highest rating",
    y = "Highest rating"
  )
```

### Who does he play with most often?
```{r echo = FALSE, message = FALSE}
# top 10 most played opponents
reg_opp <- df %>%
  group_by(Opponent) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(10)

# top 10 barchart
reg_opp %>%
  ggplot(aes(x = reorder(Opponent, n), y = n)) +
  geom_bar(stat = "identity", width = 0.6) +
  coord_flip() +
  labs(
    title = "Top 10 players he plays with most often",
    x = "Opponent",
    y = "Number of games played"
  )

# w/d/l ratio against them
df %>%
  group_by(Opponent, Result) %>%
  filter(Opponent %in% reg_opp$Opponent) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = Opponent, y = n, fill = Result)) +
  geom_bar(stat = "identity",position = "fill", width = 0.6) +
  coord_flip() +
  labs(
    title = "Results against the players he plays with most often",
    x = "Opponent",
    y = "Win:Loss:Draw"
  )
```

### which openings have higher winrates?
